Detect Pharmaceutical Quality Deviations with MS, Raman and Statistics

### 2025-02-07 dev/demo_paracetamol_fusion.Rmd
moved temporarily old vignettes and plot fixes

```{r raman-workflow}
raman_workflow <- list(
  RamanSettings_AverageSpectra_StreamFind(),
  RamanSettings_SubtractBlankSpectra_StreamFind(),
  RamanSettings_SmoothSpectra_savgol(fl = 21, forder = 4, dorder = 0),
  RamanSettings_DeleteSpectraSection_StreamFind(shiftmin = -100, shiftmax = 350),
  RamanSettings_DeleteSpectraSection_StreamFind(shiftmin = 1800, shiftmax = 2500),
  RamanSettings_CorrectSpectraBaseline_airpls(lambda = 45, differences = 1, itermax = 30),
  RamanSettings_NormalizeSpectra_minmax(),
  RamanSettings_NormalizeSpectra_blockweight()
)

@@

```{r sets-show-knn}
train_idx <- grep("_1|_2|_01|_04|_06|_10", rownames(fused_mat))

@@

```{r knn-classes}
labels <- c(
  rep("fiz", 9),
  rep("fiz_acid", 2),
  rep("fiz_ass", 2),
  rep("fiz_caff", 2),
  rep("fiz_old", 2)
)
```

@@

# MS Only

Statistical analysis is performed on MS data only.

```{r fusion}
ms_spec_mat <- ms$get_spectra_matrix()
ms_spec_mat <- mdatools::prep.autoscale(ms_spec_mat, center = TRUE, scale = FALSE)
str(ms_spec_mat)
```

```{r sets-show}
main_set_idx <- c(1:10)
ms_main_set <- ms_spec_mat[main_set_idx, ]
rownames(ms_main_set)
ms_test_set <- ms_spec_mat[-main_set_idx, ]
rownames(ms_test_set)
```

```{r pca-engine, fig.cap="MS PCA model (native plot from mdatools R package).", fig.height=9, warning=FALSE}
ms_stat <- StatisticEngine$new(analyses = ms_main_set)
ms_stat$run(StatisticSettings_MakeModel_pca_mdatools(ncomp = 3, alpha = 0.01, gamma = 0.001))
plot(ms_stat$model, showText = TRUE)
```

```{r pca-predict, warning=FALSE, fig.cap="MS PCA model prediction distances. For more information about the distances plese consult the mdatools guide in https://mdatools.com/docs/pca--distances-and-limits.html."}
ms_stat$predict(ms_test_set)
plot(ms_stat$model, showText = TRUE)
```

```{r sets-show-knn}
train_idx <- grep("_1|_2|_01|_04|_06|_10", rownames(ms_spec_mat))

ms_train_set <- ms_spec_mat[train_idx, ]
ms_train_set <- ms_train_set[order(rownames(train_set)), ]
rownames(ms_train_set)

ms_test_set <- ms_spec_mat[-train_idx, ]
rownames(ms_test_set)
```

```{r knn-classes}
labels <- c(
  rep("fiz", 9),
  rep("fiz_acid", 2),
  rep("fiz_ass", 2),
  rep("fiz_caff", 2),
  rep("fiz_old", 2)
)
```

```{r knn-engine}
ms_stat2 <- StatisticEngine$new(analyses = ms_train_set)
ms_stat2$add_classes(labels)
ms_stat2$run(StatisticSettings_MakeModel_knn(k = 3))
ms_stat2$predict(ms_test_set)
ms_stat2$model$prediction$results
```

# Raman Only

Statistical analysis is performed on Raman data only.

```{r fusion}
raman_spec_mat <- raman$get_spectra_matrix()
raman_spec_mat <- mdatools::prep.autoscale(raman_spec_mat, center = TRUE, scale = FALSE)
str(raman_spec_mat)
```

```{r sets-show}
main_set_idx <- c(1:10)
raman_main_set <- raman_spec_mat[main_set_idx, ]
rownames(raman_main_set)
raman_test_set <- raman_spec_mat[-main_set_idx, ]
rownames(raman_test_set)
```

```{r pca-engine, fig.cap="MS PCA model (native plot from mdatools R package).", fig.height=9, warning=FALSE}
raman_stat <- StatisticEngine$new(analyses = raman_main_set)
raman_stat$run(StatisticSettings_MakeModel_pca_mdatools(ncomp = 3, alpha = 0.01, gamma = 0.001))
plot(raman_stat$model, showText = TRUE)
```

```{r pca-predict, warning=FALSE, fig.cap="MS PCA model prediction distances. For more information about the distances plese consult the mdatools guide in https://mdatools.com/docs/pca--distances-and-limits.html."}
raman_stat$predict(raman_test_set)
plot(raman_stat$model, showText = TRUE)
```

```{r sets-show-knn}
train_idx <- grep("_1|_2|_01|_04|_06|_10", rownames(raman_spec_mat))

raman_train_set <- raman_spec_mat[train_idx, ]
raman_train_set <- raman_train_set[order(rownames(train_set)), ]
rownames(raman_train_set)

raman_test_set <- raman_spec_mat[-train_idx, ]
rownames(raman_test_set)
```

```{r knn-classes}
labels <- c(
  rep("fiz", 9),
  rep("fiz_acid", 2),
  rep("fiz_ass", 2),
  rep("fiz_caff", 2),
  rep("fiz_old", 2)
)
```

```{r knn-engine}
raman_stat2 <- StatisticEngine$new(analyses = raman_train_set)
raman_stat2$add_classes(labels)
raman_stat2$run(StatisticSettings_MakeModel_knn(k = 3))
raman_stat2$predict(raman_test_set)
raman_stat2$model$prediction$results
```

>>>>

```{r raman-workflow}
raman_workflow <- list(
  RamanSettings_AverageSpectra_native(),
  RamanSettings_SubtractBlankSpectra_StreamFind(),
  RamanSettings_SmoothSpectra_savgol(fl = 21, forder = 4, dorder = 0),
  RamanSettings_DeleteSpectraSection_native(min = -100, max = 350),
  RamanSettings_DeleteSpectraSection_native(min = 1800, max = 2500),
  RamanSettings_CorrectSpectraBaseline_airpls(lambda = 45, differences = 1, itermax = 30),
  RamanSettings_NormalizeSpectra_minmax(),
  RamanSettings_NormalizeSpectra_blockweight()
)

@@

```{r sets-show-knn}
train_idx <- unique(c(grep("1", rownames(fused_mat)), grep("2", rownames(fused_mat))))

@@

```{r knn-classes}
labels <- c(
  rep("fiz", 8),
  rep("fiz_acid", 2),
  rep("fiz_ass", 2),
  rep("fiz_caff", 2),
  rep("fiz_old", 2)
)
```

@@
